---
title: "**Analysis code for :**  \nEvidence for the Existence of a Bacterial Etiology for Alzheimer Disease and Evidence for a Temporal-Spatial Development of a Pathogenic Microbiome in the Brain  \nYves Moné, Joshua P. Earl, Jarosław E. Król, Azad Ahmed, Bhaswati Sen, W. Sue T. Griffin, Richard A. Jones, Sue Woodward, Garth D. Ehrlich, and Jeffrey R. Lapides"
author: "Yves Moné"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    keep_tex:  true
    #toc: yes
  html_notebook:
    depth: 3
    toc: yes
  html_document:
    keep_md: yes
    toc: yes
header-includes:
  - \usepackage{titling}
  - \pretitle{\begin{flushleft}\large\vspace*{-2.5cm}}
  - \posttitle{\end{flushleft}}  
  - \preauthor{\begin{flushleft}\large}
  - \postauthor{\end{flushleft}}  
  - \predate{\begin{flushleft}\large}
  - \postdate{\end{flushleft}}  
  - \RequirePackage{times}      # Loads the Times-Roman Fonts
  - \RequirePackage{mathptmx}   # Loads the Times-Roman Math Fonts
  - \usepackage{setspace}\onehalfspacing
  - \usepackage{indentfirst} # Indent the first line after a section header
  - \setlength{\parindent}{20pt} # Determining paragraph indentation
  #- \usepackage[left]{lineno}
  #- \linenumbers
  #- \renewcommand\linenumberfont{\normalfont\small} # Change line number font / fontsize (default: small)
  #- \setlength\linenumbersep{20pt} # Set space length between the line numbers and the text
  - \usepackage{dcolumn} # Align on the decimal point of numbers in tabular columns
  - \usepackage[labelfont=bf]{caption} # Customising captions in floating environments  (figues and tables)
  - \usepackage{float} # Improved interface for floating objects (figues and tables)
  - \usepackage{afterpage} # Execute command after the next page break
  - \usepackage{siunitx} # Comprehensive (SI) units package
  - \usepackage{amsmath} # Adjunct to serious mathematical typesetting 
geometry: margin = 0.90in
fontsize: 12pt
fig_caption: yes
indent: yes
urlcolor: blue
csl: /opt/rmarkdown_utils/plos-biology.csl
bibliography: /media/sf_ubuwin/Notebooks/SUDV_notebook/Ebola_notebook_biblio.bibtex
---
The following code was used to perform a differential abundance analysis through a Bayesian approach using the Dirichlet Multinomial Model (DMM) described in Harrison *et al.* 2020 ([doi:10.1111/1755-0998.13128](https://pubmed.ncbi.nlm.nih.gov/31872949/)).

```{r setup, eval=TRUE, include=FALSE}
knitr::opts_knit$set(
  root.dir = "/home/yves/brain_microbiome/",
  tidy.opts = list(width.cutoff=69),tidy=TRUE)

reticulate::use_python("/usr/local/bin/python3")
```

# Load packages
```{r load-packages, eval=FALSE}
# Microbiome analysis
library("phyloseq")
library("decontam")
library("zCompositions")
library("CoDaSeq")

# Graphics
library("ggplot2")
library("ggrepel")
library("grid")
library("gridExtra")
library("ggpubr")
library("lemon")

```

```{r variable, eval=FALSE}
# Set working directory
workDir = "brain_microbiome"
dir.create(file.path(getwd(), workDir))
setwd(file.path(getwd(), workDir))

# Create subdirectories for data and results
dir.create(file.path(getwd(), "data"), showWarnings = FALSE)
dir.create(file.path(getwd(), "results"), showWarnings = FALSE)
```

# User-defined Functions
```{r user-fxn, eval=TRUE}
# Identify contaminants (Prevalence method)
## ps is a phyloseq object
isNotContam <- function(ps){
  sample_data(ps)$is.neg <- sample_data(ps)$sample_type == "NegCtrl"
  nc_05 <- isNotContaminant(ps,
    neg="is.neg",
    threshold = 0.5,
    normalize = TRUE, detailed = TRUE, method = "prevalence")
  return(nc_05)
}

# Create a composite feature of infrequent OTU across samples
# ps: phyloseq object
# prev: prevalence threshold (percentage)
# rab: relative abundance threshold (percentage)
# other: A logical.Default TRUE. The function returns the filtered
#        phyloseq object with a composite feature named "OTU_others".
#        If FALSE, then the function returns the filtered
#        phyloseq object and the list of OTU that have been removed.

filter.rare <- function(ps, prev, rab, other = TRUE){
  prev = 20
  rab = 0.005
  ps_prev<-filter_taxa(ps, 
                      function(x) sum(x>0)<=((prev/100)*length(x)),
                      prune = FALSE)
  ps_rab <- transform_sample_counts(ps,
                                   function(x){round(x/sum(x),6)})
  ps_ab <- filter_taxa(ps_rab,
                       function(x) mean(x) <= (rab/100),
                       prune = FALSE)
  fTaxa <- ps_prev & ps_ab
  rmtaxa <- taxa_names(ps)[fTaxa]
  if (other == TRUE) {
    ps.new <- merge_taxa(ps, eqtaxa=rmtaxa, archetype=1)
    taxa_names(ps.new)[taxa_names(ps.new) == rmtaxa[1]] <- "OTU_others"
    return(list(ps.new, rmtaxa))
  } else {
    keepTaxa <- !fTaxa
    ps.new <- prune_taxa(keepTaxa, ps)
    return(list(ps.new, rmtaxa))
  }
}

# Process DMM results
## If 95% of the samples of the distribution of the difference 
dmm.sig <- function(res_diff){
  # res_diff = differences in pi parameters
  sig <- vector()
  no_sig <- vector()
  for(i in 1:dim(res_diff)[2]){
    perc <- length(which(res_diff[,i] > 0 ))/ length(res_diff[,i])
    if(perc >= 0.95 | perc <= 0.05){
      sig <- c(sig, i)
    }else{no_sig <- c(no_sig, i)
  }
}
return(list(significant = sig,
  nosignificant = no_sig))
}

## Plot DMM results
plot.dmm <- function(comparison, otuFile, pi1.file, pi2.file,
  otu2tax, res_dmm_sig, out_tsv, OTU_all=FALSE){

  # comparison = vector of conditions to compare. eg: c(AD, control)
  # otuFile: OTU table (tab-separated text file)
  # pi1.file: estimated pi parameters in AD group
  # pi2.file: estimated pi parameters in control group
  # otu2tax: species name assigned to each OTU
  # res_dmm_sig: output list from dmm.sig function
  # out_tsv: Table of bacteria that shift in relative abundance
  # OTU_all: TRUE to display all the OTU names in the plot,
  #          FALSE to display the name of differentially abundant OTU

  # Input files
  otu_df <- read.table(otuFile,sep="\t",row.names=1,h=T)
  pi1_df <- read.table(pi1.file,sep="\t",h=T)
  pi2_df <- read.table(pi2.file,sep="\t",h=T)
  otu2sp_df <- read.table(otu2tax,sep="\t",h=T)

  otu_table <- otu_df[, grep(pattern="OTU_", colnames(otu_df))]

  # Differences in estimated pi parameters
  diffs <- pi1_df - pi2_df

  # Mean of differences in pi parameters
  mean_diffs <- data.frame("mean"=apply(diffs, 2, FUN = mean))
  mean_diffs$cilow <- apply(diffs, 2,
    FUN = quantile, probs = c(0.025))
  mean_diffs$ciup <- apply(diffs, 2,
    FUN = quantile, probs = c(0.975))
  mean_diffs$otu_id <- colnames(otu_table)
  idx_sig <- colnames(otu_table[,(res_dmm_sig$significant),
                                drop=FALSE])
  mean_diffs$dab <- mean_diffs$otu_id %in% idx_sig

  # Prepare data to plot
  plot_data <- mean_diffs[order(mean_diffs[,1]),]
  plot_data$otu_idx <- c(1:dim(plot_data)[1])
  otu_id_lvl <- plot_data$otu_id[order(plot_data$mean)]
  plot_data$otu_id <- factor(plot_data$otu_id,
                             levels = otu_id_lvl)
  plot_data$otu_sig <- plot_data$otu_id[ifelse(plot_data$dab==TRUE,
                                               plot_data$otu_id,NA),
                                        drop=FALSE]

  # Table of differentailly abundant OTU
  dab_tab <- merge(na.omit(plot_data[,c(1,2,7,8)]), otu2sp_df,
                   by.x="otu_sig", by.y="OTU_ID", all.x=TRUE)
  names(dab_tab) <- c("OTU_id",
                      "mean.diff",
                      "Index",
                      "Phylum",
                      "Species")
  dab_df <- data.frame(dab_tab[,c(3,1,4,5,1,2)],
                       "Species"=apply(dab_tab[5], 2,
                                       function(x)
                                       as.character(gsub("_",
                                                         " ",
                                                         x))),
                       row.names = NULL)
  dab_df2 <- dab_df[order(dab_df$Index),]
  write.table(dab_df2,
              paste0("results/",out_tsv),
              sep="\t",row.names=FALSE, quote=FALSE)

  # Select OTU to be labeled on the plot
  if(OTU_all == FALSE){
    otu_labs <- plot_data[na.omit(plot_data$otu_sig),]
  }else{
    otu_labs <- plot_data
  }
  otu_labs <- merge(otu_labs, otu2sp_df,
                    by.x="otu_sig", by.y="OTU_ID", all.x=TRUE)
  otu_labs$annot <- ifelse(is.na(otu_labs$species),
                           paste0(otu_labs$otu_sig),
                           paste0(otu_labs$otu_sig,"-",
                                  otu_labs$species))
  otu_labs <- otu_labs[order(otu_labs$mean),]
  otu_labs <- otu_labs[,c(2:dim(otu_labs)[2],1)]

  # Select label position
  mynudge.x <- ifelse(otu_labs[,1] < 0, 25, -0.002)
  mynudge.y <- ifelse(otu_labs[,1] < 0, -0.009, 0.009)
  myhjust <- ifelse(otu_labs[,1] < 0, 0, 1.8)

  plot_list <- list()

  # Mean_diff plot
  p1 <- ggplot(plot_data) +
    geom_point(aes(x=otu_idx, y=mean,
                   color=relevel(factor(dab), "TRUE")),
               size = 1.5) +
    geom_errorbar(aes(x=otu_idx, y=mean,
                      color=relevel(factor(dab), "TRUE")),
                  ymin = plot_data$cilow, ymax = plot_data$ciup) +
    geom_hline(yintercept = 0.00, color="grey28", linetype = 2) +
    geom_text_repel(data= otu_labs,
                    aes(x=otu_idx,y=mean, label=annot,
                        color=relevel(factor(dab), "TRUE")),
                    vjust=0.5, hjust=myhjust,
                    size=4,
                    fontface="bold",
                    colour = "black",
                    nudge_x = mynudge.x,
                    nudge_y = mynudge.y,
                    segment.color = "black",
                    segment.alpha = 0.5,
                    segment.size = 0.5
                    )+
    ylab("Effect size
         (difference in estimated relative abundance)") +
    ylim(c(min(plot_data$cilow)*1.1,
           max(plot_data$ciup)*1.5)) +
    coord_capped_cart(bottom ='both') +
    scale_x_continuous(
      breaks = c(seq(0, max(plot_data$otu_idx),by=20),
                 max(plot_data$otu_idx)),
      labels = c(seq(0, max(plot_data$otu_idx),by=20),""),
      expand = expansion(mult=c(0.05,0.016)),
      name="Index of OTU (ordered by effect size)") +
    annotate("text",
             x=dim(plot_data)[1]-25,
             y=0.05,
             label= paste(comparison[1]),
             colour = "black", fontface = "bold", size = 5) +
    annotate("text",
             x=dim(plot_data)[1]-25,
             y=-0.025,
             label= paste(comparison[2]),
             colour = "black", fontface = "bold", size = 5) +
    theme_bw() +
    theme(legend.position="none",
          axis.line.x = element_line(colour="black", linetype = 1),
          axis.ticks.x = element_line(color="black"),
          axis.text.x = element_text(size=14, color="black"),
          axis.title.x = element_text(size=14, color="black"),
          axis.line.y = element_line(colour="black", linetype = 1),
          axis.title.y = element_text(size=14, color="black"),
          axis.text.y=element_text(size=14, color="black"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.border = element_blank(),
          plot.margin = unit(c(1,2,1,2), "cm")
          )
  plot_list[[1]] <- p1
  plot_list[[2]] <- plot_data
  return(plot_list)
}
```

# Load dataset
```{r dataset, eval=FALSE}
ps <- readRDS("data/brain_phyloseq.rds")
```
# Detection of contaminant OTU and data filtering

Potential contaminant OTUs were detected based on their mean relative abundance and their prevalence in experimental samples vs. negative controls using the R package Decontam (Davis *et al.* 2018, [doi:10.1186/s40168-018-0605-2](https://pubmed.ncbi.nlm.nih.gov/30558668/)). To qualify as contaminant, an OTU had to have a score $\geq$ 0.5 or a higher mean relative abundance in the negative controls than the biological samples.

```{r rab-contam, eval=FALSE}
# Mean relative abundance (RAb) in negative controls
ps_neg <- subset_samples(ps, sample_type == "NegCtrl")
pneg <- prune_taxa(taxa_sums(ps_neg) > 0, ps_neg)
pneg_rab<-transform_sample_counts(pneg,
                                  function(x){round(x / sum(x),6)})
neg_mean_rab <- apply(X = otu_table(pneg_rab),
                      MARGIN = ifelse(taxa_are_rows(pneg_rab),
                                      yes = 1, no = 2),
                      FUN = function(x){mean(x)})

# Mean RAb in experimental samples
ps_true <- subset_samples(ps, sample_type == "sample")
exp_rab<-transform_sample_counts(ps_true,
                                  function(x){round(x / sum(x),6)})
exp_mean_rab <- apply(X = otu_table(exp_rab),
                      MARGIN = ifelse(taxa_are_rows(exp_rab),
                                      yes = 1, no = 2),
                      FUN = function(x){mean(x)})

# Compare mean RAb between negative control and experimental samples
otu_exp_neg<-round(exp_mean_rab[which(names(exp_mean_rab) %in%
                                      names(neg_mean_rab))]*100, 4)
rab_df <- data.frame("OTU_ID" = names(neg_mean_rab),
                     "rab.neg" = round(neg_mean_rab*100,4),
                     "rab.sample" = otu_exp_neg)
rab_df$greater.in.ctrl <- ifelse(rab_df[,2] > rab_df[,3],
                                 TRUE, FALSE)

# decontam package - Prevalence-based method
## conf_lvl: confidence levels from suppl. table S11
tax_df <- as.data.frame(tax_table(ps))
otu2tax <- data.frame("OTU_ID"= rownames(tax_df),
                      "phylum"=tax_df$phylum,
                      "species"=tax_df$species)
conf_lvl <- read.table("data/brain_confidence.tsv", sep="\t", h=T)
otu_neg1 <- as.data.frame(otu_table(pneg))
otu_neg1$OTU_ID <- rownames(otu_neg1)
otu_neg <- merge(otu_neg1, conf_lvl[, c(1,7:8)], by="OTU_ID")
otu_tax_neg <- merge(otu_neg, otu2tax, by="OTU_ID")
otu_tax_neg <- otu_tax_neg[order(otu_tax_neg$OTU_ID),]

## Identification of contaminants
non_contam <- isNotContam(ps)
contam <- non_contam[which(!is.na(non_contam$p) &
                           non_contam$not.contaminant==FALSE),]
contam$OTU_ID <- rownames(contam)
contam_df <- merge(otu_tax_neg, contam, by="OTU_ID", all.x=TRUE)
write.table(contam_df,
            "results/OTU_decontamPrev.tsv",
            sep="\t", row.names=FALSE, quote=FALSE)

## Combine results from RelAb and decontam approaches
df1 <- merge(rab_df, contam, by="OTU_ID", all.x=TRUE)
df2 <- merge(otu_tax_neg[, c(1,6,7,9)], df1[,c(1:4,9,10)],
             by="OTU_ID", all.x=TRUE)

contam <- df2[order(df2$species),]
names(contam) <- c("OTU_ID",
                       "genus_conf",
                       "species_conf",
                       "species",
                       "RAb.NegCtrl",
                       "RAb.samples",
                       "RAb.contaminant",
                       "decontam.p_value",
                       "decontam.not.contaminant")
contam$contaminant <- ifelse(contam$RAb.contaminant==TRUE,
                                 TRUE,
                                 ifelse(!(is.na(contam[,8])),
                                        TRUE, FALSE))
write.table(contam,
            "results/OTU_ContamDetection.tsv",
            sep="\t",row.names=FALSE,quote=FALSE)

```

```{r filter-data, eval=FALSE}
#
### Filter-out contaminant OTU
contam <- read.table("results/OTU_ContamDetection.tsv",
                     sep="\t",h=T)

rmOTU <- contam[which(contam$contaminant==TRUE),1]
ncOTU <- !(taxa_names(ps) %in% rmOTU)
ps_nc <- prune_taxa(ncOTU, ps)

# Remove negative control samples
ctrl <- c("UA_C1_2_lbc47", "UA_C2_2_lbc62",
          "NW_C1_lbc37", "NW_C2_lbc38",
          "UA_C1_1_lbc68", "UA_C2_1_lbc76")
ps_nc2 <- subset_samples(ps_nc, !(sample_names(ps_nc) %in% ctrl))

# Remove low yield samples (cut-off=100 reads)
ps100 <- prune_samples(sample_sums(ps_nc2)>=100, ps_nc2)

# Remove samples with 0 counts across all OTU
ps_s0 <- prune_samples(sample_sums(ps100) > 0, ps100)

# Remove OTU with 0 counts across all samples
ps_decontam <- prune_taxa(taxa_sums(ps_s0) > 0, ps_s0)

## Infrequent OTU are grouped into a composite feature "OTU_others"
### OTU_others: RAb cut-off = 0.005% and prevalence cut-off = 20%
ps_filtered <- filter.rare(ps_decontam, 20, 0.005, other = TRUE)
saveRDS(ps_filtered[[1]], "results/brain_phyloseq_filtered.rds")

other <- taxa_names(ps_decontam) %in% ps_filtered[[2]]
ps_other <- prune_taxa(other, ps_decontam)

write.table(tax_table(ps_other),
            "results/list_otu_other.tsv",
            sep="\t",row.names=FALSE,quote=FALSE)
#

```

# Differential abundance testing

```{r dmm-input, eval=FALSE}
# Prepare data for DMM analysis
ps <- readRDS("results/brain_phyloseq_filtered.rds")
otu_df <- as.data.frame(t(as.matrix(otu_table(ps))))
meta_df <- data.frame(sample_data(ps))

dmm_df <- merge(meta_df[,c(5:7)], otu_df, by=0)
colnames(dmm_df)[1] <- "sample_ID"
dmm_df <- dmm_df[order(dmm_df$diagnosis, dmm_df$specimen_ID),]
write.table(dmm_df,
  "results/dmm_brain_table.tsv",
  sep="\t",row.names=FALSE,quote=FALSE)

```

\newpage
## Model specification
The Dirichlet Multinomial model was specified in the Stan probabilistic programming language.
```{stan stan-dmm, output.var="model", eval=FALSE}
// DM model
data {
  int<lower=1> notus; // total counts in each sample
  int<lower=1> nreps; // number of replicates
  int<lower=1> N; // number of sampling groups (AD vs. C)
  int<lower=1> M; // number of specimens (individual biopsies)
  int<lower=1> start[N]; // group sampling start indices
  int<lower=1> end[N]; // group sampling end indices
  int<lower=1> reptospe[nreps]; // index of replicate per specimen
  int datamatrix[nreps, notus];
}

parameters {
  real<lower=0> theta[N];
  real<lower=0> tau[M];
  simplex[notus] psi[M];
  simplex[notus] pi[N];
  simplex[notus] p[nreps];
}

model {
  for(m in 1:M){
    target += exponential_lpdf(tau[m] | 0.01);
    target += dirichlet_lpdf(psi[m] | rep_vector(0.0000001, notus));
  }
  for(i in 1:N){
    for(j in start[i]:end[i]){
      target += exponential_lpdf(theta[i] | 0.01);
      target += dirichlet_lpdf(pi[i] | tau[reptospe[j]]*psi[reptospe[j]]);    
      target += dirichlet_lpdf(p[j] | theta[i]*pi[i]);
      target += multinomial_lpmf(datamatrix[j,] | p[j]);
    }
  }
}

```

\newpage
## Run the Dirchlet Multionmial model
The Dirichlet Multinomial model was run through Python using the Pystan package. DMM is computationally intensive and have been run on a high performance cluster. 
```{python, eval=FALSE}
import os, sys
import pickle
import time
import pystan
import pandas as pd
import numpy as np
#
n_cpu=os.cpu_count()
print("n_cpu:", n_cpu)
print("Python:", sys.version)
print("pystan:", pystan.__version__)
print("pandas:", pd.__version__)
print("numpy:", np.__version__)
#
method = 'HMC'
modelfile = "data/DMM_brain.stan"
count_table = "results/dmm_brain_table.tsv"
prefix = "ADvsC"
pathname = "results"
# Output directory
os.makedirs('{0}/DMM-{2}_output_{1}'.format(pathname,prefix,method),
            exist_ok=True)
# Load data
otu_table = pd.read_csv(count_table,sep='\t')
otu_dat = otu_table.sort_values(by=['diagnosis', 
          'specimen_ID']).reset_index(drop=True)
#
# get group indices (Stan language uses 1-based indexing)
starts = [1,otu_dat['diagnosis'].value_counts()[0]+1]
ends = [otu_dat['diagnosis'].value_counts()[0],
        len(otu_dat['diagnosis'])]
groups = [starts,ends]
#
# get specimen indices (Stan language uses 1-based indexing)
codes, uniques = pd.factorize(otu_dat['specimen_ID'])
otu_dat['reptospe'] = codes
reptospe = list(codes +1)
#
# Data matrix
counts = otu_dat.filter(regex='OTU_.*',axis=1)
#
#
data = {
	'datamatrix': counts+1,
    'nreps': int(len(counts)),
    'notus': int(len(counts.columns)),
    'N': int(len(starts)),
    'M': int(len(uniques)),
    'start': groups[0],
    'end': groups[1],
    'reptospe': reptospe
    }
#
start_time = time.process_time()
#
sm = pystan.StanModel(file=modelfile, model_name='DMM')
fit = sm.sampling(data=data,
	chains=4,
	control = {'max_treedepth': 15},
	warmup = 1500,
	iter = 3500,
	algorithm = "NUTS",
	seed = 123,
	thin = 2,
	pars = ['pi','psi'],
	verbose = True,
	sample_file =
	'{0}/DMM-{2}_output_{1}/DMM_{2}_{1}_sample_file.csv'.format(
	  pathname,prefix,method)
	)
# Save object fit
with open('{0}/DMM-{2}_output_{1}/DMM_fit_{2}_{1}.pkl'.format(
  pathname,prefix,method),'wb') as f:
    pickle.dump({'model' : sm, 'fit' : fit}, f, protocol=-1)
#
fit_df = fit.to_dataframe()
fit_df.to_csv('{0}/DMM-{2}_output_{1}/DMM_fit_{2}_{1}.tsv'.format(
  pathname,prefix,method), sep='\t', header=True, index=False)
#
# Extract pi parameters for each sampling group
est_pi = fit.extract(pars='pi')
mat_pi1 = np.asmatrix(est_pi['pi'][:,0,:])
mat_pi2 = np.asmatrix(est_pi['pi'][:,1,:])
pi1_df = pd.DataFrame(mat_pi1, )
pi1_df.to_csv('{0}/DMM-{2}_output_{1}/DMM_{2}_{1}_est_pi1.tsv'.format(
  pathname,prefix,method), sep='\t', header=True, index=False)
pi2_df = pd.DataFrame(mat_pi2)
pi2_df.to_csv('{0}/DMM-{2}_output_{1}/DMM_{2}_{1}_est_pi2.tsv'.format(
  pathname,prefix,method), sep='\t', header=True, index=False)
#
# Summary
with open('{0}/DMM-{2}_output_{1}/DMM_fit_{2}_{1}_summary.txt'.format(
  pathname,prefix,method),'w') as f:
	print(fit.stansummary(), file = f)
#
# Time elapsed
end_time = time.process_time()
time_elapsed = end_time - start_time
print("Time elapsed: ", time_elapsed, "seconds")

```

\newpage
## Processing DMM results
```{r res-dmm, eval=FALSE}
# Generate otu2tax table
ps <- readRDS("results/brain_phyloseq_filtered.rds")
tax_df <- as.data.frame(tax_table(ps))
otu2sp <- data.frame("OTU_ID"= rownames(tax_df),
                      "phylum"=tax_df$phylum,
                      "species"=tax_df$species)
write.table(otu2sp, 
            "results/brain_otu2sp.tsv",
            sep="\t",row.names=FALSE,quote=FALSE)

comp <- c("AD", "C")
otu_file = "results/dmm_brain_table.tsv"
pi1 = "results/DMM_HMC_ADvsAMC_est_pi1.tsv"
pi2 = "results/DMM_HMC_ADvsAMC_est_pi2.tsv"
otu2tax = "results/brain_otu2sp.tsv"
out_tsv = "results/DMM_results.tsv"

# OTU differentially abundant
pi1_df <- read.table(pi1,sep="\t",h=T)
pi2_df <- read.table(pi2,sep="\t",h=T)
diffs <- pi1_df - pi2_df
res_dmm_sig <- dmm.sig(diffs)
length(res_dmm_sig$sig)
#
## Plot differences

plot_others <- plot.dmm(
  comp,
  otuFile = otu_file,
  pi1.file = pi1,
  pi2.file = pi2,
  otu2tax,
  res_dmm_sig,
  out_tsv,
  OTU_all = FALSE)
#
ggsave(plot = plot_others[[1]],
       "results/Differential_abundance_plot.tiff"),
       device="tiff", width = 38, height = 30,
       units = 'cm', dpi = 400, compression = "lzw")

```

